---
title: "<center> **How to use ClustersAnalysis Package**</center><br/>"
author: ""
date: "<center>`r format(Sys.time(), '%d %B %Y')`</center><br/><br/><br/>Table of contents"
output:
  html_document:
    toc: yes
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warnings=FALSE)
```

## Introduction

This is a demonstration of using the R package Clusters. You will see how to analyze classes according to one or more variables. The group variable must be of the type factor or character and the exploratory variables can be quantitative or qualitative. In this demonstration we are going to use natives dataset from R such as "iris", "infert" or "esoph".

### Short Descriptions of datasets

#### Iris

The data set consists of 50 samples from each of three species of Iris (Iris setosa, Iris virginica and Iris versicolor). Four features were measured from each sample: the length and the width of the sepals and petals, in centimeters. Based on the combination of these four features, Fisher developed a linear discriminant model to distinguish the species from each other.

```{r iris}
summary(iris)
```

#### Infert
This is a matched case-control study dating from before the availability of conditional logistic regression.
```{r infert}
summary(infert)
```

### Esoph
Data from a case-control study of (o)esophageal cancer in Ille-et-Vilaine, France. This is a data frame with records for 88 age/alcohol/tobacco combinations.
```{r esoph}
summary(esoph)
```


## Import ClustersAnalysis from Github (using devtools)


```{r}
#Use the below line to install devtools if necessary
#install.packages("devtools")
library(devtools)

#install package from github
devtools::install_github("clepadellec/ClustersAnalysis")

#load package
library(ClustersAnalysis)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
ClustersAnalysis::u_plot_size_effect((Univariate_object(infert,1)))
```

