---
title: '**How to use ClustersAnalysis Package**'
output:
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warnings=FALSE)
```

# Introduction

This is a demonstration of using the R package ClustersAnalysis. You will see how to analyze classes according to one or more variables. The group variable must be of the type factor or character and the exploratory variables can be quantitative or qualitative. In this demonstration we are going to use natives dataset from R such as "iris", "infert" or "esoph".

## Short Descriptions of datasets

### Iris : 

The data set consists of 50 samples from each of three species of Iris (Iris setosa, Iris virginica and Iris versicolor). Four features were measured from each sample: the length and the width of the sepals and petals, in centimeters. Based on the combination of these four features, Fisher developed a linear discriminant model to distinguish the species from each other.

```{r iris}
summary(iris)
```

### Infert : 
This is a matched case-control study dating from before the availability of conditional logistic regression.
```{r infert}
summary(infert)
```

### Esoph : 
Data from a case-control study of (o)esophageal cancer in Ille-et-Vilaine, France. This is a data frame with records for 88 age/alcohol/tobacco combinations.
```{r esoph}
summary(esoph)
```


# Import ClustersAnalysis from Github (using devtools)



```{r}
#Use the below line to install devtools if necessary
#install.packages("devtools")
#library(devtools)

#install package from github
#Use the below line to install ClustersAnalysis if necessary
#devtools::install_github("clepadellec/ClustersAnalysis")

#load package
library(ClustersAnalysis)

```


## How to access to help

You can juste use the fonction help(function name) to see all the documentation about your function.
```{r}
help("u_plot_size_effect")
```


# Univariate Analysis for qualitatives variariables

To begin we will try to understand, for each qualitative explanatory variable, if it affects the group variable. It's necessary to create an object of univariate type. You can use the constructor **Univariate_object**.
```{r}

#Creation of univariate object using esoph dataframe and "agegp" (first column) as the group variable
u_esoph<-Univariate_object(esoph,1)

#detail of attributes associated with the object
print(u_esoph)
```

## Contingency table and size effect

The first thing we can do is to create a contingency table between the explanatory variable and the group variable and then visualize lines/columns profils. In our case the explanatory variable is "tobgp" which is the tobacco consumption (gm/day). To do this you can use the **u_desc_profils**
```{r}
#use intecract=TRUE to show an interactive graphique with widgets like zoom, comparisons...
ClustersAnalysis::u_desc_profils(u_esoph,3,interact=FALSE)
```
The distributions don't show any particular phenomena. The most represented classes are the 35-44  and 55-64 years. Then we can see that there are more than a half that smokes less than 20 g/days. The only fact that we can see is that the 75+ people are close to 75% to don't smoke a lot.


Now we are going to see in details if there is a size effect between these two variables. To do this we can use **u_desc_size_effect** which return the test statistic vt (comparison between proportions). Then we can also use **u_plot_size_effect** which create a mosaic plot.

```{r}
u_desc_size_effect(u_esoph,3)
#use intecract=TRUE to show an interactive graphique with widgets like zoom, comparisons...
u_plot_size_effect(u_esoph,3,interact=FALSE)
```
If we refer to the results we can see that the biggest "vt" values are for 75+ peoples who smokes 0-9g/day or 10-19 g/day. So this is for these two combinations that one can most easily conclude that there is an over-representation. We can use the mosaic plot to confirm our comment.

You can do the sames analyses for the variable "tobgp".


## Chisq test

We can now use **u_chisq_test_all** which is use to calculate the p-avlue for a chisq test between the group variable and each qualitatives variables.

```{r}
u_chisq_test_all(u_esoph)
```
Here we can see that the p-avlues are very high which mean that we accept the null hypothesis and the variables are independant. Because of this, it's useless to calculate v-cramer.

If you want to see significant results you can run the code below, the aim is to predict the sexe according to many qualitatives attributs. We can see that for each chisq.test, we can conclude that the sex depends on the attributs. But the attributs which has the most influence is "qualif".
```{r}
#install.packages("questionr")
#library(questionr)
# data(hdv2003)
# d<- hdv2003[,3:8]
# u_chisq_test_all(Univariate_object(d,1))
```

## Correspondence analysis

To better visualize if some groups are close to a particular modality we can use correspondence analysis. The **u_afc_plot** function return a biplot to visualize our results.
```{r}
u_afc_plot((Univariate_object(esoph,1)),3)
```

In our case the principal information that we can see is that the 25-34 peoples are quite close to 30+g/day. It's a new information that we didn’t have before. 


# Univariate Analysis for quantitatives variables

Now we are going to use the famous iris dataset. The aim si to understand how the Species can be described.
```{r}

#Creation of univariate object with the "Species" as group variable
u_iris <-Univariate_object(iris,5)
```


## Means comparisons (using student test)

To begin we are going to search for each group and each explanatory variable if the fact to belong to a species or not have an influence on the length and the width of Petal/Sepal. To do this we use **u_ttest_all**, this function for each group, separates the group variable into two modalities : target modality or others modality. Then we verify normality hypothesis and if it's possible we use a student test to define if it's significative or not.
```{r}
 u_ttest_all(u_iris)
```
Here we can see that each group have one or many significatives variables which mean that the fact to belong to setosa,versicolor or virginica have an influence on the attributs.

## Fisher Test
```{r}
u_fisher_test_all(u_iris)
```


## Test Value for clustering (for only one explanatory variable)
```{r}
d<-iris[,c(1,5)]

u_test_value(Univariate_object(d,2))

```

## Silhouettes indices
```{r}
u_silhouette_ind(u_iris)
u_silhouette_plot(u_iris, interact=FALSE)
```



## Principal component analysis
```{r}
u_sil_pca_plot(u_iris,interact = FALSE)
```

## Rand Index
```{r}
u_kmean_rand_index(u_iris)
```

# Multivariate Analysis
